### 2019-10-07 makefile.Rmd
some comments starting the process of gettin this paper ready for Ecosphere

---
title: "extinction risk makefile, compiliation notes"
author: "Daniel Reuman"
date: ""
output: pdf_document
---

<!--checkpoint package-->
```{r checkpoint_chunk, echo=F, warning=F, message=F, results="hide"}
library(checkpoint)
if (!dir.exists("./.checkpoint/")){
  dir.create("./.checkpoint/")
}
checkpoint("2019-01-13",checkpointLocation = "./")
```

# Get tex files and pdf files for the supp mat, but the pdf files 
#are missing the external references

```{r tex_for_supp_mat}
rmarkdown::render(input="SuppMat.Rmd") 
```

# Get tex files and pdf files for the main text, but the pdf files 
#are missing the external references

```{r tex_for_main_text}
rmarkdown::render(input="MainText.Rmd") 
```

# Now recompile the latex

```{r recompile_latex}
#do it once
system2(command="pdflatex",args="SuppMat.tex",stdout=TRUE,stderr=TRUE)
system2(command="pdflatex",args="MainText.tex",stdout=TRUE,stderr=TRUE)

#do it again to make sure to get the external references right
system2(command="pdflatex",args="SuppMat.tex",stdout=TRUE,stderr=TRUE)
system2(command="pdflatex",args="MainText.tex",stdout=TRUE,stderr=TRUE)
```